---
output: github_document
---


[![Travis-CI Build Status](https://travis-ci.org/r-gris/rangl.svg?branch=master)](https://travis-ci.org/r-gris/rangl) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/r-gris/rangl?branch=master&svg=true)](https://ci.appveyor.com/project/r-gris/rangl) [![Coverage Status](https://img.shields.io/codecov/c/github/r-gris/rangl/master.svg)](https://codecov.io/github/r-gris/rangl?branch=master)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme-figure/README-"
)
```

## Tidy tables for spatial data 

The basics for creating a mesh from a `SpatialPolygons` object is in place. 

Multiple multi-part objects are decomposed to a set of related, linked tables. Object identity is maintained with attribute metadata and this is carried through to colour aesthetics in 3D plots.  


```{r,eval=FALSE}
library(rglwidget)
library(rgl)
library(rworldmap)
data(countriesLow)


library(rangl)
cmesh <- tri_mesh(countriesLow[sample(nrow(countriesLow), 10), ])
plot(cmesh)

#subid <- currentSubscene3d()
#rglwidget(elementId="plot_flat")
```

```{r,eval=FALSE,include=FALSE}
library(rangl)
#ogrListLayers("inst/extdata/water_lake_superior_basin.gpkg")
spdf <- rgdal::readOGR("inst/extdata/water_lake_superior_basin.gpkg", "lake_superior_basin")

lake <- tri_mesh(spdf)

plot(lake)
```
Particular todos: 

- systematize the structure vs. relation indexes, right now it's a bit brute force
- default plotting aesthetics use the `viridis` palette
- the halo radius is not right . . . 


## Build notes

* Travis needs libproj-dev installed, via .travis.yml (CRAN does not)
* R-hub needs SystemRequirements: PROJ.4
* All need export RGL_USE_NULL=TRUE
* plot in examples is fine, but there's a warning from R CMD check in interactive mode (??)

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

