---
title: "Examples with anglr"
author: "Michael D. Sumner"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
vignette: >
  %\VignetteIndexEntry{Examples with anglr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE, results="asis"}
source("setup.R")
knitr::opts_chunk$set(rgl.newwindow = TRUE)
set.seed(123)
```

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>")
rgl::par3d(windowRect = c(100, 100, 512 + 100, 512 +100))
library(raster)
library(maptools)
```

# Introduction

The 'anglr' package illustrates some generalizations of GIS-y and topology tasks in R with "tables". See the [package project](https://github.com/r-gris/anglr/blob/master/README.md) for more information. 


Get some maps and plot in 3D - in plane view, or globe view. 

**NOTE:** these plots are interactive, but performance and quality will be better (for now) if run locally. 




```{r,eval=TRUE, webgl=TRUE}

library(rgl)
library(maptools)
data(wrld_simpl)
library(raster)
library(sf)
## convert to triangles and plot
library(anglr)
cmesh <- anglr(st_as_sf(wrld_simpl))
plot(cmesh)
rglwidget()
```



```{r, webgl=TRUE}
sids <- raster::shapefile(system.file("shapes/sids.shp", package="maptools"))
projection(sids) <- "+proj=longlat +ellps=clrk66"

ex <- extent(sids) + 5
gl <- graticule::graticule(seq(xmin(ex), xmax(ex), length = 15), 
                           seq(ymin(ex), ymax(ex), length = 8))


## convert to triangles, but wrap onto globe then plot
smesh <- anglr(sids)
plot(globe(smesh))
mgl <- anglr(gl)
mgl$o$color_ <- "black"
plot(globe(mgl), lwd = 2)
```



## Holes are (trivially) supported. 

It's trivial to have "holes", because there are no holes, because we have a true surface, composed of 2D primitives (triangles). 

```{r,eval=TRUE, webgl=TRUE}
library(spbabel)
data(holey)

## SpatialPolygonsDataFrame
sph <- sp(holey)

glh <- anglr(sph)
plot(glh)
```




## Lines are supported. 


```{r, webgl=TRUE}
linehouse <- as(sph, "SpatialLinesDataFrame")
plot(anglr(linehouse))
```




## Globe lines

```{r, webgl=TRUE}
lmesh <- anglr(as(wrld_simpl, "SpatialLinesDataFrame"))
plot(globe(lmesh))
```





## rgl objects

Rgl mesh3d objects that use "triangle" primitives are supported. 

```{r,webgl=TRUE}
library(rgl)

dod <- anglr(dodecahedron3d(col = "cyan"))
octo <- anglr(translate3d(octahedron3d(col = "blue"), 6, 0, 0))
plot(dod, col = viridis::viridis(5)[1], alpha = 0.3)
plot(octo, col = viridis::viridis(5)[5], alpha = 0.3)
bg3d("grey")
```



To complete the support for these rgl objects we need quads, and to allows a mix of quads and triangles in one data set (that's what `extrude3d` uses). Extrusion is a bit limiting so it's unclear how useful this is (yes it is common, though). See `rgl::extrude3d` for the most readily available method in R. 


## Points


And points work!  (Don't laugh). 


```{r, webgl=TRUE}
library(anglr)
library(maptools)
data(wrld_simpl)
mpts <- as(as(wrld_simpl, "SpatialLinesDataFrame"), "SpatialMultiPointsDataFrame")
plot(anglr(mpts))
rgl::view3d(theta = 25, phi = 3)
```


## Trips

The trip package includes  a 'walrus818' data set courtesy of Anthony Fischbach.  Zoom around and see if you can find them. 

```{r, webgl=TRUE}
library(trip)
library(anglr)
data(walrus818)

library(graticule)
prj <-"+proj=laea +lon_0=0 +lat_0=90 +ellps=WGS84"
gr <- graticule(lats = seq(40, 85, by = 5), ylim = c(35, 89.5), proj = prj)
library(maptools)
data(wrld_simpl)

w <- spTransform(subset(wrld_simpl, coordinates(wrld_simpl)[,2] > -70),  prj)
library(graticule)
walrus <- spTransform(walrus818, prj)

gr$color_ <- "black"
rgl::par3d(windowRect = c(100, 100, 912 + 100, 912 +100))
plot(anglr(gr))
w$color_ <- sample(viridis::inferno(nrow(w)))
plot(anglr(w), specular = "black")
plot(anglr(walrus))
um <- structure(c(0.934230506420135, 0.343760699033737, 0.0950899347662926, 
                  0, -0.302941381931305, 0.905495941638947, -0.297159105539322, 
                  0, -0.188255190849304, 0.24880850315094, 0.950081348419189, 0, 
                  0, 0, 0, 1), .Dim = c(4L, 4L))
par3d(userMatrix = um)

```

